---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---


## Loading and preprocessing the data

I use the packages dplyr, lubridate and ggplot2 and also lattice. 

```{r echo=FALSE}
require(dplyr)
require(lubridate)
require(ggplot2)
require(lattice)
```
Loading the file is done with:

```{r echo=TRUE}
activity<-read.csv("./activity.csv",sep=",",dec=".",header = TRUE)
date<-ymd(activity$date)
activity<-as.data.frame(activity)
```

I use mutate do make new day variable with wday function from lubridate.
```{r echo=TRUE}

act<-mutate(activity,day=wday(date,label=FALSE))

```
## What is mean total number of steps taken per day?

The number of steps taken per day is: 
```{r echo=TRUE}


tot<-aggregate(steps~day,act,sum)
print(tot,type="html")

```
And the total is: 
```{r echo=TRUE}
sum(tot, na.rm=TRUE)

```

I use the ggplot function to make the histogram of the total number of steps taken each day is: 
```{r echo=TRUE}
h<-aggregate(steps~date,act,sum)

step_hist<-ggplot(h,aes(steps))
step_hist+geom_histogram(fill="darkgreen")+xlab("Steps")+labs(title="Histogram of total number of steps\n")



```

The mean and median of the total number of steps taken per day is:

```{r echo=TRUE}
mean(h$steps,na.rm=TRUE)
median(h$steps,na.rm=TRUE)
```


## What is the average daily activity pattern?



The time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r echo=TRUE}
avg_steps<-aggregate(steps~interval,act,mean)
time_avg<-ggplot(avg_steps, aes(interval,steps))
time_avg+geom_line()+labs(title="Average number of steps\n")+labs(x="Interval")+labs(y="Average Steps")
```

Which 5-minutes interval, on average acorss all the days in the dataset, 
contains the maximum of steps:
```{r echo=TRUE}
loc<-which.max(avg_steps$steps)
avg_steps[loc,1,]

```
And the maximum average number of steps is: 
```{r echo=TRUE}
avg_steps[loc,2,]
```



## Imputing missing values

The total number of missing values in the dataset is:

```{r echo=TRUE}
v<-is.na(act$step)
sum(v)

```
You need to define a strategy. This example is where we take the average of the day in case of a NA. 
```{r echo=TRUE}
meanday<-mean(h$steps,na.rm=TRUE)
c<-rep(0,times=nrow(act))
for (i in 1:nrow(act)){ 
    if (is.na(act[i,1,])==TRUE){c[i]<-meanday} else {c[i]<-act[i,1]}
}
c<-as.numeric(c)
```

The new dataset with the imputed values is: 
```{r echo=TRUE}
new_data<-cbind(c,act)
colnames(new_data)<-c("steps_imputed","steps","date","interval","day")
```

The histogram of the total number of steps taken each day: 
```{r echo=TRUE}
v<-aggregate(steps_imputed~date,new_data,sum)

step_hist<-ggplot(v,aes(steps_imputed))
step_hist+geom_histogram(fill="darkgreen")+xlab("Steps")+labs(title="Histogram of total number of steps\n")

```

Regarding the appearance of the histogram above this is because that on 1.october 2012 there was no observations and thereby the histogram will have two bars. The mean and median is: 

```{r echo=TRUE}
mean(v$steps_imputed)
median(v$steps_imputed)

```

The estimates differ from the first part.The reason for this is that on the 1. october 2012 there was no observations the entire day and thereby the imputation will overwrite NA to the mean from the first day. 


## Are there differences in activity patterns between weekdays and weekends?

Definition of a new variable. I use the mutate function. I have already made a variable day with use of the wday function (wday equals 1 on sunday and 7 on saturday).

```{r echo=TRUE}
act_new<-mutate(new_data,weekday=(ifelse((day==1)|(day==7),"weekend","weekday")))

```
first I find total number of steps for each interval and for each type of day (weekday/weekend):
```{r echo=TRUE}
total_steps1<-aggregate(steps_imputed~interval+weekday,act_new,mean)
```
For making the histogram I use the lattice plotning system:


```{r echo=TRUE}
xyplot(steps_imputed~interval|weekday,type="l",data=total_steps1,layout=c(1,2),xlab="Interval", ylab="Number of steps")
```
